NIHAO VII: Predictions for the galactic baryon budget in dwarf to Milky Way mass haloes

### 2016-04-22 ms_nihao_budget.tex
revised budget paper

\begin{abstract}
  We use the NIHAO galaxy formation simulations to make predictions
  for the baryonic budget  in present day galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses.  The sample is made of 88 independent high resolution
  cosmological zoom-in simulations.  NIHAO galaxies reproduce key
  properties of observed galaxies, such as the stellar mass vs halo
  mass and cold gas vs stellar mass relations. Thus they make
  plausible predictions for the baryon budget.  We present the mass
  fractions of stars, cold gas ($T<10^4$K), cool gas ($10^4 < T <
  10^5$K), warm-hot gas ($10^5 < T < 10^7$K), and hot gas ($10^7$K$ <
  T$), inside the virial radius, $R_{200}$.  Compared to the predicted
  baryon mass, using the dark halo mass and the universal baryon
  fraction, $f_{\rm b}\equiv \Omega_{\rm b}/\Omega_{\rm m}=0.15$, we
  find that for all of our haloes, the missing mass is not just outside 
  the virial radius $\Rvir$, it has been relocated past 2$\Rvir$.  
  Haloes of mass $\Mhalo\sim 10^{10}\Msun$ are missing $\sim
  90\%$ of their baryons.  More massive haloes ($\Mhalo\sim
  10^{12}\Msun$) retain a higher fraction of their baryons, with $\sim
  30\%$ missing, consistent with observational estimates.    
%  Finally we show that the baryonic budget of most massive 
%  galaxies in NIHAO suite can match the observation for most 
%  components but the cool gas, which requires further 
%  theoretical and observational work to find the answer where
%  the cool gas actually is.
\end{abstract}

@@

%% missing baryon problem
On cosmological scales, the ratio between the total baryonic and
gravitating mass is measured to be $f_{\rm b}\equiv \Omega_{\rm
  b}/\Omega_{\rm m}\simeq 0.16$ (The Planck Collaboration 2014).
However, the cold baryonic mass density implied by several galaxy baryon
estimates is only 3-8\% of the big bang nucleosynthesis expection
\citep{Persic92, Fukugita98,  Bell03, McGaugh10}.  The majority of the
cosmic baryons are thought to be in the form of hot gas around or
between galaxies \citep{Cen09}. Until recently only a fraction of
these baryons had been detected \citep{Bregman07, Shull12}.  This
discrepancy is referred to as the ``missing baryon problem''.
Recent advances in the detection of gas in the circum
galactic medium (CGM) have come from the COS survey \citep{Tumlinson11,
  Tumlinson13, Thom12,  Werk12, Werk13}.  On the scale of Milky Way
mass haloes $\Mhalo \sim 10^{12}\Msun$ a significant amount of warm
($10^4 < T < 10^7$K) gas has been detected \citep{Werk14}, accounting
for 33-88\% of the baryon budget. In the future such observations will
be extended to a wider range of halo masses.

@@

@@

We define the fiducial baryonic mass as:
\begin{equation}
M_{\rm b} \equiv M_{\rm b}(R_{200})= \frac{f_{\rm b}}{1-f_{\rm b}}M_{\rm dm}(R_{200}) 
\end{equation}
where the $M_{\rm dm}$ is the total dark matter mass in the halo, and
the $f_{\rm b} = \Omegab/\Omegam \sim 0.15$ is the cosmic baryon
fraction (the ratio between baryon density and mass density including baryonic mass
plus dark matter), so that the  $M_{\rm b}$ is the baryonic mass inside the
virial radius if the baryons follow the dark matter closely.

@@

In Fig.~\ref{fig:corona}, we present the radial distribution of 
gas in different phases normalized to the

@@

%% FIGURE 4
\begin{figure}
\centerline{
  \psfig{figure=budget.eps,width=0.5\textwidth}
}
\caption{Baryonic budget (the fraction of different components) 
  of $3.5\times 10^{11} < M_{200}/\Msun <
  3.5 \times 10^{12}$ NIHAO galaxies (blue points 
  with 1$\sigma$ error bars) compared with observations of 
  M$_{200}$ = 10$^{12.2}$ $\Msun$ galaxies (shaded regions)
  }
\label{fig:comparison}
\end{figure}

@@

The observations and the simulations match well in every phase
outside of the cool CGM gas, where the observations find $3\times$
the mass that simulations predict.  If the observations are correct,
the simulations have either ejected too far gas that could be cool, or 
they have created a CGM with the wrong mix of gas temperatures. The
total gas fractions (0.39 in COS-HALOs, 0.41 in NIHAO) suggest the
latter option.  As the CGM of lower mass galaxies will soon be observed ,
Table~\ref{tab:comperison} lists information about CGM
mass fractions of the different components of gas in haloes down to a
halo mass of $\sim 10^{10}\Msun$.

@@

@@

@@

\item For the highest mass haloes in our study $\sim 10^{12}\Msun$ 
      our simulations are consistent with the observed fractions of stars, 
      cold gas, warm and hot gas. However, we underpredict 
      the cool ($10^4 < T < 10^5 K$) gas with $f_{\rm cool}=0.11\pm0.06$ 
      vs 0.28-0.48 observed.

\item This discrepancy may indicate that our feedback model pushes the 
      gas too far, or puts gas in the wrong phases, 
      and can be resolved with future observations. 
      Since we match the overall gas fraction we think that  
      there might be some problem in mixing the gas temperature.
\end{itemize}

@@

@@

@@

@@

@@

@@

@@

>>>>

\begin{abstract}
  We use the NIHAO galaxy formation simulations to make predictions
  for the baryonic budget  in present day galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses.  The sample is made of 88 independent high resolution
  cosmological zoom-in simulations.  NIHAO galaxies reproduce key
  properties of observed galaxies, such as the stellar mass vs halo
  mass and cold gas vs stellar mass relations. Thus they make
  plausible predictions for the baryon budget.  We present the mass
  fractions of stars, cold gas ($T<10^4$K), cool gas ($10^4 < T <
  10^5$K), warm-hot gas ($10^5 < T < 10^7$K), and hot gas ($10^7$K$ <
  T$), inside the virial radius, $R_{200}$.  Compared to the predicted
  baryon mass, using the dark halo mass and the universal baryon
  fraction, $f_{\rm b}\equiv \Omega_{\rm b}/\Omega_{\rm m}=0.15$, we
  find that for all of our haloes, the missing mass is not just outside 
  the virial radius $\Rvir$, it has been relocated past 2$\Rvir$.  
  Haloes of mass $\Mhalo\sim 10^{10}\Msun$ are missing $\sim
  90\%$ of their baryons. 
  {\bf Cool gas dominates the fraction of total baryons surrounding
  galaxies up to several times of virial radius, in which most of
  them are well outside of virial radius.}
   More massive haloes ($\Mhalo\sim
  10^{12}\Msun$) retain a higher fraction of their baryons, with $\sim
  30\%$ missing, consistent with observational estimates.    
%  Finally we show that the baryonic budget of most massive 
%  galaxies in NIHAO suite can match the observation for most 
%  components but the cool gas, which requires further 
%  theoretical and observational work to find the answer where
%  the cool gas actually is.
\end{abstract}

@@

%% missing baryon problem
On cosmological scales, the ratio between the total baryonic and
gravitating mass is measured to be $f_{\rm b}\equiv \Omega_{\rm
  b}/\Omega_{\rm m}\simeq 0.16$ (The Planck Collaboration 2014).
However, the cold baryonic mass density implied by several galaxy baryon
estimates is only 3-8\% of the big bang nucleosynthesis expection
\citep{Persic92, Fukugita98,  Bell03, McGaugh10}.  The majority of the
cosmic baryons are thought to be in the form of hot gas around or
between galaxies \citep{Cen09}. Until recently only a fraction of
these baryons had been detected \citep{Bregman07, Shull12}.  This
discrepancy is referred to as the ``missing baryon problem''.
{\bf Several theoretical research with cosmological simulation
have constrained the phase status and cosmological enviroments of
the potential reservior to hide the missing baryons 
\citep{Yoshida05, He05, Dave10, Zhu11, Haider16}.}

{\bf The circum galactic medium (CGM) is one major potential 
reservior of missing baryons in halo scale.}
Recent advances in the detection of  gas in the CGM have come 
from the COS survey \citep{Tumlinson11,
  Tumlinson13, Thom12,  Werk12, Werk13}.  On the scale of Milky Way
mass haloes $\Mhalo \sim 10^{12}\Msun$ a significant amount of warm
($10^4 < T < 10^7$K) gas has been detected \citep{Werk14}, accounting
for 33-88\% of the baryon budget. In the future such observations will
be extended to a wider range of halo masses.
{\bf To understand the physical properties of CGM and the origins 
of components in each phase has been shown to be able to test the
impact of feedback models \citep{Sharma12,Marasco13}.
\citet{Dave09} predicted galactic halo baryon fractions of
galaxies with halo mass range from $10^{11} \Msun$ to
$10^{13} \Msun$
cosmological hydrodynamic simulations with a well-constrained
model for galactic outflow, which found that, without outflow
model, the baryon fraction is roughly the cosmic baryonic 
fraction, but with outflow in the simulation, the baryon fraction
is increasingly suppressed in haloes to lower mass. By comparing
results at z=3 and z=0, they showed the large haloes remove their
baryons at early epoch while small haloes lose baryons more
recently due to wind material takes longer to return to low-mass
galaxies than high-mass galaxies.
\citet{Sokolowska16} studied halo gas of three Milky way-sized 
galaxies using cosmological zoom-in simulations. They found that
most of missing baryons actually resides in warm-hot and hot gas
which contribute to 80\% of the total gas reservior. 
The recovered baryon fraction within 3 virial radii is 90\% and 
can solve the missing baryon problem.
The warm-hot medium is sensitive to the feedback model so that
the reliable spatial mapping of warm-hot medium will provide 
a stringent test for feedback models.}

@@

%% FIGURE 4
\begin{figure}
\centerline{
  \psfig{figure=missing_vr.eps,width=0.5\textwidth}
}
\caption{Baryon radius as function of total mass within virial 
         radius. The points are color coded by the ratio 
         between baryon radius to virial radius of each galaxy.}
\label{fig:missvr}
\end{figure}

%% FIGURE 5
\begin{figure*}
\centerline{
  \psfig{figure=budget_inout.eps,width=0.99\textwidth}
}
\caption{Mass fraction of gas mass in four phases to the fiducial
         baryonic mass within virial radius. The blue points are
         mass fraction for the gas inside virial radius and the 
         red points are gas between virial radius and baryon
         radius.}
\label{fig:inout}
\end{figure*}

@@

We define the fiducial baryonic mass as:
\begin{equation}
M_{\rm b} \equiv M_{\rm b}(R_{200})= \frac{f_{\rm b}}{1-f_{\rm b}}M_{\rm dm}(R_{200}) 
\label{equ:mb}
\end{equation}
where the $M_{\rm dm}$ is the total dark matter mass in the halo, and
the $f_{\rm b} = \Omegab/\Omegam \sim 0.15$ is the cosmic baryon
fraction (the ratio between baryon density and mass density including baryonic mass
plus dark matter), so that the  $M_{\rm b}$ is the baryonic mass inside the
virial radius if the baryons follow the dark matter closely.

@@

{\bf In Fig.~\ref{fig:corona}, we present the radial distribution of 
gas in different phases normalized to the total baryon mass in
corresponding radial bins at z=0.
All simulations share a common attribute.
The cold gas (T$< 10^4$ K) in corona mostly locates 
in very center region 
($R < 0.5 R_{\rm vir}$) where most stars in galaxies form.
The cool ($10^4$K $<$ T $<$ $10^5$K) and 
warm-hot ($10^5$K $<$ T $<$ $10^7$K) gas expand to large 
distance contrastly,
the fractions keep constant till 2 times $R_{\rm vir}$.
The hot gas (T$>10^7$K) is minority in corona for 
all galaxies in NIHAO
sample, the maximum of the fraction of hot gas is less
than 5\%.}

{\bf Despite these similarities, we find considerably high proportion
of cool gas for lower mass galaxies (M$_{\rm tot} < 10^{11} \Msun$)
in the whole corona region.
For higher mass galaxies, warm-hot gas dominates the corona 
which signals higher efficiency  of feedback.
Even beyond virial radius, the cool and warm-hot gas has
similar features as the gas within virial radius which reveals
the gas surrounding galaxies within large distance is the major
reservior of baryons.}

@@

{\bf To find out how far the baryons escape, we measured the baryon 
radius within which the total baryon mass equals to the fiducial
baryonic mass defined by Equ.~\ref{equ:mb}.
Fig.~\ref{fig:missvr} illustrates the baryon radius of each galaxy
as function of the total mass within virial mass.
We find the baryon radius increases as the total mass goes higher
generally.
In the whole mass range, The baryon radii normalized by 
corresponding virial radii span a similar range
which are from larger than virial radius to more than 5 times
virial radius, except there are three galaxies with baryon
radii larger than 8 times virial radius. 
It shows that the normalized distance where the baryons inside
virial radius can reach is not sensitive to the halo masses.}

{\bf In Fig.~\ref{fig:inout}, we calculate the fractions of 
these four kinds of gas inside virial radius and between 
virial radius and baryon radius. The upper-left panel shows the 
fractions of cold gas. The fractions of cold gas between
virial radius and baryon radius of
galaxies with halo masses in the range between $10^{9} \Msun$ to 
$10^{10} \Msun$ are greater than 80\%. As the halo mass increase,
this fractions go down dramatically, in the mass range between
$10^{10} \Msun$ and $10^{11} \Msun$ the fractions are less than 
20\% for most of galaxies and the fractions of galaxies with halo
mass above $10^{11} \Msun$ are almost 0. 
For the cold gas inside virial radius, the fractions increase
gradually as halo masses increase and the maximum fraction is
about 20\% with halo mass at $10^{12} \Msun$.
The upper-right panel is about the fractions of cool gas. 
The fractions of cool gas in region outside of virial radius 
increase from 0 at halo masses are $10^9 \Msun$ to 80\% at halo 
masses are around $10^{10} \Msun$ to $10^{11} \Msun$, then 
decrease in higher mass range and reach 0 again at halo masses
are $10^{12} \Msun$. The cool gas insdie virial radius shares 
a similar trend but the maximum fraction is about 40\% at 
halo masses are $10^{11} \Msun$.
The fractions for warm-hot gas are shown in the lower-left panel,
the trends of the gas inside and outside gas are quite similar
which increase as the halo masses increase monotonically and whose 
maximum values are only 20\%.
The hot gas shown in the lower-right panel is absolutely minority,
the fraction of gas inside baryon radius is almost 0 in the 
whole halo mass range.
We thus conclude that, for the galaxies with halo mass 
below $10^{11}$, the majority of baryons associated with the
halo mass within virial radius are in the cold and cool phases, 
meanwhile are well outside of virial radius, for the gas inside
virial radius, the majority is cool gas for the galaxies with
halo masses are $10^{11} \Msun$ and when halo masses are higher
than $10^{11} \Msun$ the fractions of cold gas, cool gas and 
warm-hot gas are comparable.}

%% FIGURE 6
\begin{figure}
\centerline{
  \psfig{figure=budget.eps,width=0.5\textwidth}
}
\caption{Baryonic budget (the fraction of different components) 
  of $3.5\times 10^{11} < M_{200}/\Msun <
  3.5 \times 10^{12}$ NIHAO galaxies (blue points 
  with 1$\sigma$ error bars) compared with observations of 
  M$_{200}$ = 10$^{12.2}$ $\Msun$ galaxies (shaded regions)
  }
\label{fig:comparison}
\end{figure}

@@

The observations and the simulations match well in every phase
outside of the cool CGM gas, where the observations find $3\times$
the mass that simulations predict.  If the observations are correct,
the simulations have either ejected too far gas that could be cool, or 
they have created a CGM with the wrong mix of gas temperatures. The
total gas fractions (0.39 in COS-HALOs, 0.41 in NIHAO) suggest the
latter option.  
{\bf However, \citet{Stern16} developed a new method to constrain the
physical condtions in the cool CGM from measurements of ionic 
columns denisities. This new method combined the information
available from different during the photoionization modeling,
and was applied to the COS-HALOs data, 
yielding a total cool CGM mass within the virial radius
of $1.3\times10^{10}\Msun$ which is shown by the green hashed
bar in Fig~\ref{fig:budget} and have good agreement with
our prediction.
As the Fig.~\ref{fig:corona} and Fig.~\ref{fig:inout} show, the
cool gas is the most important component in CGM so that
the more correct knowledge about physical properties of CGM are
necessary to understand the role the CGM on galaxy formation
better.}
As the CGM of lower mass galaxies will soon be observed ,
Table~\ref{tab:comperison} lists information about CGM
mass fractions of the different components of gas in haloes down to a
halo mass of $\sim 10^{10}\Msun$.

@@

\item {\bf Cold gas in most of galaxies (except several low mass
      galaxies with low SFR) locates within 0.5 times virial 
      radius. The cool gas dominates the corona at low mass range
      end and the fraction of warm-hot gas increase with 
      the increasing of halo mass and become majority at high
      mass end.}

@@

\item {\bf Cool gas is major component in total baryons 
      whose mass is same as the fiducial baryonic mass of 
      total dark matter mass inside virial radius. Most of the
      cool gas in these baryons locates out of the virial radius.}

@@

\item For the highest mass haloes in our study $\sim 10^{12}\Msun$ 
      our simulations are consistent with the observed 
      fractions of stars, 
      cold gas, warm and hot gas.

\item We underpredict 
      the cool ($10^4 < T < 10^5 K$) gas 
      with $f_{\rm cool}=0.11\pm0.06$ 
      vs 0.28-0.48 observed of COS-HALOs.
      {\bf \citet{Stern16} employed new method which have
      a tight constraint on CGM properties shows a 
      similar result as we predict.
      To solve this discrepancy may be helpful to answer
      if current feedback model pushes the 
      gas too far, or puts gas in the wrong phases, 
      and can be resolved with future observations.}

\end{itemize}

@@

% Missing Halo Baryons and Galactic Outflows
\bibitem[Dave{\'e} (2009)]{Dave09} Dav{\'e}, R.\ 2009, ASPC, 419, 347D

% The intergalactic medium over the last 10 billion years - I. Lya absorption and physical conditions
\bibitem[Dav{\'e} et al.(2010)]{Dave10} Dav{\'e}, R., Oppenheimer, B.~D., Katz, N., et al.\ 2010, \mnras, 408, 2051

@@

% Large-scale mass distribution in the Illustris simulation
\bibitem[Haider et al.(2016)]{Haider16}
Haider, M., Steinhauser, D., Vogelsberger, M., et al.\ 2016, \mnras, 457, 3024

% Distributions of the Baryon Fraction on Large Scales in the Universe
\bibitem[He et al.(2005)]{He05}
He, P., Feng, L.~L., Fang, L.~Z.\ 2005, \apj, 623, 601

@@

% On the origin of the warm-hot absorbers in the Milky Way's halo
\bibitem[Marasco et al.(2013)]{Marasco13} 
Marasco, A., Marinacci, F., Fraternali, F.\ 2013, \mnras, 433, 1634

@@

% On the structure of hot gas in haloes: implications for the LX-TX relation and missing baryons
\bibitem[Sharma et al.(2012)]{Sharma12}
Sharma, P., McCourt, M., Parrish, I.~J., Quataert, E.\ 2012, \mnras, 427, 1219

@@

% Diffuse Coronae in Cosmological Simulations of Milky Way-sized Galaxies
\bibitem[Sokolowska et al.(2016)]{Sokolowska16}
Sokolowska, A., Mayer, L., Babul, A., Madau, P., Shen, S.\ 2016, \apj, 819, 21

% A Universal Density Structure for Circum-Galactic Gas
\bibitem[Stern et al.(2016)]{Stern16} 
Stern, J., Hennawi, J.~F., Prochaska, J.~X., Werk, J.~K.\ 2016, arXiv:1604.02168

@@

% The Temperature Structure of the Warm-Hot Intergalactic Medium
\bibitem[Yoshida et al.(2005)]{Yoshida05} 
Yoshida, N., Furlanetto, S.~R., Hernquist, L.\ 2005, \apj, 618L, 91

@@

% Dynamical effect of the turbulence of the intergalactic medium on the baryon fraction distribution
\bibitem[Zhu et al.(2011)]{Zhu11} Zhu, W., Feng, L.~L., Fang, L.~Z.\ 2011, \mnras, 415, 1093
